A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}

### 2022-04-13 ICLP Paper/samplepaper.tex
small changes

The task of optimizing the horizon of a (partial) schedule means choosing
the execution order of operations of the latest time window sharing a machine
such that the value for the \lstinline{makespan} variable is minimized.
In single-shot ASP modulo DL solving with \clingodl, this can be accomplished
via the command-line option \lstinline{--minimize-variable=makespan}.
\hl{This option, however, is implemented by means of a fixed control loop that
cannot be combined with (other) multi-shot solving methods.}
For the successive optimization by time windows, where the scheduling
horizon gradually increases, we thus require a dedicated treatment of DL constraints
limiting the value for \lstinline{makespan}.
To this end, the \lstinline{optimize(m)} subprogram below line~\ref{prg:diff_log:oper_min}
declares an external atom \lstinline{horizon(m)} for controlling whether a DL
constraint asserted in line~\ref{prg:diff_log:oper_min:end} is active and limits
the \lstinline{makespan} value to an integer supplied for the parameter~\lstinline{m}.

@@

\hl{The main steps of our control loop for successive schedule optimization
by multi-shot ASP modulo DL solving, implemented by means of the Python interface of \clingodl,
are displayed in Figure~{\ref{fig:enum}} and further detailed in the following.}
When launching the optimization process for a new time window, any instance of the
\lstinline{horizon(m)} atom introduced before is set to false in step~1 for making sure that
some (partial) schedule is feasible.
\hl{No such atoms have been introduced yet for the first time window $w=1$,
where the static (default) subprogram called} \lstinline{base}\hl{,
supplying a JSP instance along with its decomposition in terms of facts over} \lstinline{window/3}\hl{,
and the} \lstinline{step(}$1$\lstinline{)} \hl{subprogram for operations of the first time window
are instantiated in steps~2 and~3.}
%
\begin{figure}[tb]
  \begin{tcolorbox}[breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
  %  \vspace{0.1\baselineskip}
  Starting with $H:=\emptyset$, for $w:= 1$ to number $n$ of time windows:
  \begin{enumerate}[label=\textbf{\arabic*.}]
    \item For each $h\in H$, set \lstinline{horizon(}$h$\lstinline{)} to false
    \item % \begin{tabbing}
          If $w= 1$, % \= 
          then $P:=\{\text{\lstinline{base}}\}$;
          % \\
          % \>

          % \hfill
          else $P := \{\text{\lstinline{start((}$j$\lstinline{,}$s$\lstinline{),}$t$\lstinline{,}$w-1$\lstinline{).}}
                      \mid
                      \{\text{\lstinline{window(}$j$\lstinline{,}$s$\lstinline{,}$w-1$\lstinline{)}},
                      \text{\lstinline{(}$j$\lstinline{,}$s$\lstinline{)}}=t\}\subseteq X\}$
          % \end{tabbing}
    \item Ground $P\cup\{\text{\lstinline{step(}$w$\lstinline{)}}\}$
    \item While there is some answer set $X$ and the time limit per time window is not reached:
      \begin{enumerate}[label=\textbf{\alph*.}]
      \item $h:=t-1$, where $($\lstinline{makespan}${}=t)\in X$
      \item If $h\notin H$, then $H := H\cup\{h\}$ and ground $\{$\lstinline{optimize(}$h$\lstinline{)}$\}$
      \item Set \lstinline{horizon(}$h$\lstinline{)} to true
      \end{enumerate}
    \end{enumerate}
  \end{tcolorbox}
  \caption{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
  \end{figure}%
  % \captionof{figure}{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
%
Once some schedule with a horizon $h+1$ is found in step~4a,
the step~4b consists of instantiating
the subprogram \lstinline{optimize(}$h$\lstinline{)} % is instantiated
on demand, i.e.,
in case $h$ has not been passed as a value for~\lstinline{m} before, and
the corresponding \lstinline{horizon(}$h$\lstinline{)} atom is set to true in step~4c for activating the DL
constraint reducing the admitted scheduling horizon to~$h$.
The step 4 of successively reducing the horizon~$h$ in order to find better partial schedules
% is repeated until 
stops when the imposed \lstinline{makespan} value turns out to be infeasible,
meaning that an optimal partial schedule has been found.
As already mentioned above, the introduced instances of the external \lstinline{horizon(m)} atom
are then set to false in step~1, and the successive optimization proceeds by in steps~2 and~3 instantiating the
\lstinline{step(}$w+\nolinebreak1$\lstinline{)} subprogram for the next time window $w+1$ (if any) and
also supplying the determined starting times of operations from time window~$w$
by reified facts.
The described control loop for successively extending good-quality partial schedules to
a global solution 
% is implemented by means of the Python interface of \clingodl,
% where a time limit 
can be configured with a time limit,
included as secondary stopping condition in step~4,
to restrict the optimization efforts per time window
and thus make sure that the iterative scheduling progresses.

>>>>

The task of optimizing the horizon of a (partial) schedule means choosing
the execution order of operations of the latest time window sharing a machine
such that the value for the \lstinline{makespan} variable is minimized.
In single-shot ASP modulo DL solving with \clingodl, the minimization can be accomplished via the command-line option \lstinline{--minimize-variable=makespan}.
\hl{This option, however, is implemented by means of a fixed control loop that
cannot be combined with (other) multi-shot solving methods.}
For the successive optimization by time windows, where the scheduling
horizon gradually increases, we thus require a dedicated treatment of DL constraints
limiting the value for \lstinline{makespan}.
To this end, the \lstinline{optimize(m)} subprogram below line~\ref{prg:diff_log:oper_min}
declares an external atom \lstinline{horizon(m)} for controlling whether a DL
constraint asserted in line~\ref{prg:diff_log:oper_min:end} is active and limits
the \lstinline{makespan} value to an integer supplied for the parameter~\lstinline{m}.

@@

\hl{The main steps of our control loop for successive schedule optimization
by multi-shot ASP modulo DL solving, implemented by means of the Python interface of \clingodl,
are displayed in Figure~{\ref{fig:enum}} and further detailed in the following.}
When launching the optimization process for a new time window, any instance of the
\lstinline{horizon(m)} atom introduced before is set to false in step~1 for making sure that
some (partial) schedule $X$ is feasible.
\hl{No such atoms have been introduced yet for the first time window $w=1$,
where the static (default) subprogram called} \lstinline{base}\hl{,
supplying a JSP instance along with its decomposition in terms of facts over} \lstinline{window/3}\hl{,
and the} \lstinline{step(}$1$\lstinline{)} \hl{subprogram for operations of the first time window
are instantiated in steps~2 and~3.}
%
\begin{figure}[tb]
  \begin{tcolorbox}[breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
  %  \vspace{0.1\baselineskip}
  Starting with $H:=\emptyset$, for $w:= 1$ to number $n$ of time windows:
  \begin{enumerate}[label=\textbf{\arabic*.}]
    \item For each $h\in H$, set \lstinline{horizon(}$h$\lstinline{)} to false
    \item % \begin{tabbing}
          If $w= 1$, % \= 
          then $P:=\{\text{\lstinline{base}}\}$;
          % \\
          % \>

          % \hfill
          else $P := \{\text{\lstinline{start((}$j$\lstinline{,}$s$\lstinline{),}$t$\lstinline{,}$w-1$\lstinline{).}}
                      \mid
                      \{\text{\lstinline{window(}$j$\lstinline{,}$s$\lstinline{,}$w-1$\lstinline{)}},
                      \text{\lstinline{(}$j$\lstinline{,}$s$\lstinline{)}}=t\}\subseteq X\}$
          % \end{tabbing}
    \item Ground $P\cup\{\text{\lstinline{step(}$w$\lstinline{)}}\}$
    \item While there is some answer set $X$ and the time limit per time window is not reached:
      \begin{enumerate}[label=\textbf{\alph*.}]
      \item $h:=t-1$, where $($\lstinline{makespan}${}=t)\in X$
      \item If $h\notin H$, then $H := H\cup\{h\}$ and ground $\{$\lstinline{optimize(}$h$\lstinline{)}$\}$
      \item Set \lstinline{horizon(}$h$\lstinline{)} to true
      \end{enumerate}
    \end{enumerate}
  \end{tcolorbox}
  \caption{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
  \end{figure}%
  % \captionof{figure}{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
%
Once some schedule $X$ with a horizon $h+1$ is found in step~4a,
the step~4b consists of instantiating
the subprogram \lstinline{optimize(}$h$\lstinline{)} % is instantiated
on demand, i.e.,
in case $h$ has not been passed as a value for~\lstinline{m} before, and
the corresponding \lstinline{horizon(}$h$\lstinline{)} atom is set to true in step~4c for activating the DL
constraint reducing the admitted scheduling horizon to~$h$.
The step 4 of successively reducing the horizon~$h$ in order to find better partial schedules
% is repeated until 
stops when the imposed \lstinline{makespan} value turns out to be infeasible,
meaning that an optimal partial schedule has been found.
As already mentioned above, the introduced instances of the external \lstinline{horizon(m)} atom
are then set to false in step~1, and the successive optimization proceeds by in steps~2 and~3 instantiating the
\lstinline{step(}$w+\nolinebreak1$\lstinline{)} subprogram for the next time window $w+1$ (if any) and
also supplying the determined starting times of operations from time window~$w$
by reified facts.
The described control loop for successively extending good-quality partial schedules to
a global solution 
% is implemented by means of the Python interface of \clingodl,
% where a time limit 
can be configured with a time limit,
included as secondary stopping condition in step~4,
to restrict the optimization efforts per time window
and thus make sure that the iterative scheduling progresses.
